project-plan

### 2025-09-26 project-plan.qmd

Previous studies and reviews have discussed preventable trauma deaths where recurring patterns occur, namely, delays in imaging and care, missed injuries, deviations in care regarding airway-related care, haemorrhage control, and fluid management (12–16). Importantly, some of these findings were observed in high-performing level I trauma centres, underscoring that even mature systems remain vulnerable to error. Furthermore, injuries to the central nervous system account for a substantial proportion of trauma-related deaths, as demonstrated in analyses from large registries such as the Arkansas Trauma Registry (16). Yet, while previous research has described OFI in trauma resuscitation and in general ICU trauma cohorts, relatively little is known about the prevalence, nature, and determinants of OFI specifically among patients with TBI. <!-- Add a part after ACS TQIP using WHO as a reference to TQIP-->

Previous papers at Karolinska University Hospital have contributed to this field by describing OFIs across different
aspects of trauma care. In 2023, Hussein A. et al. analysed patient- and process-related factors during trauma
resuscitation and concluded that variables such as time to first CT and injury severity were strongly associated with
OFI (17). Cao ES. et al. continued the focus on ICU cohorts, presenting that 9.9% ICU-admitted adult trauma patients
experienced at least one OFI (1). Most recently, Szolnoky K et al. assessed OFI patterns in trauma cohorts and
demonstrated that, although overall incidence declined over time, patients with TBI remained likely to have
preventable shortcomings in care (18). <!-- Rewrite into a more general rationale paragraph-->

@@

For continuous variables, medians and interquartile ranges (25th-75th  percentiles) are used. Categorical variables
are presented using the number of patients and their corresponding percentages of the cohort. <!-- If multivariable regression analysis will be used, how will we handle missing data from other variables in the same patient? -->

>>>>

Previous studies and reviews have discussed preventable trauma deaths where recurring patterns occur, namely, delays in imaging and care, missed injuries, deviations in care regarding airway-related care, haemorrhage control, and fluid management (12–16). Importantly, some of these findings were observed in high-performing level I trauma centres, underscoring that even mature systems remain vulnerable to error. Furthermore, injuries to the central nervous system account for a substantial proportion of trauma-related deaths, as demonstrated in analyses from large registries such as the Arkansas Trauma Registry (16). Yet, while previous research has described OFI in trauma resuscitation and in general ICU trauma cohorts, relatively little is known about the prevalence, nature, and determinants of OFI specifically among patients with TBI.

TBI remains a challenge within modern trauma systems, not only due to its contribution to mortality but also due to
the magnifying impact on the outcome because of small process deviations. Delays in imaging, inconsistent airway
management, and lapses in haemodynamic measures are all examples that can result in irreversible damage. Although
TQIP have been developed to identify and correct such lapses, with the WHO stressing the importance of linking OFIs
to corrective actions and verification of bettered outcomes, most research has been done on mixed trauma cohorts.
This, on the other hand, leaves TBI-specific vulnerabilities undercharacterised. Recent work by Szolnoky et al.
suggests that, despite overall declines in OFI, patients with TBI remain at risk of preventable shortcomings,
underscoring the need for TBI-specific quality improvement efforts (23).

<!--Previous papers at Karolinska University Hospital have contributed to this field by describing OFIs across different
aspects of trauma care. In 2023, Hussein A. et al. analysed patient- and process-related factors during trauma
resuscitation and concluded that variables such as time to first CT and injury severity were strongly associated with
OFI (17). Cao ES. et al. continued the focus on ICU cohorts, presenting that 9.9% ICU-admitted adult trauma patients
experienced at least one OFI (1). Most recently, Szolnoky K et al. assessed OFI patterns in trauma cohorts and
demonstrated that, although overall incidence declined over time, patients with TBI remained likely to have
preventable shortcomings in care (18).-->

@@

For continuous variables, medians and interquartile ranges (25th-75th  percentiles) are used. Categorical variables
are presented using the number of patients and their corresponding percentages of the cohort. <!-- If multivariable regression analysis will be used, how will we handle missing data from other variables for the same patient? -->
